\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-12-01 papers/jss/article.Rnw
Rewrite mapreduce introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@921 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Protocol Buffers have been used extensively at Google for almost all
RPC protocols, and for storing structured information in a variety of
persistent storage systems since 2000 \citep{dean2009designs}.  The
\pkg{RProtoBuf} package has been in widespread use by hundreds of
statisticians and software engineers at Google since 2010.  This
section describes a simplified example of a common design pattern of
collecting a large structured data set in one language for later
analysis in \proglang{R}.

Many large data sets in fields such as particle physics and information
processing are stored in binned or histogram form in order to reduce
the data storage requirements \citep{scott2009multivariate}.  In the
last decade, the MapReduce programming model \citep{dean2008mapreduce}
has emerged as a popular design pattern that enables the processing of
very large data sets on large compute clusters.

Many types of data analysis over large data sets may involve very rare
phenomenon or deal with highly skewed data sets or inflexible
raw data storage systems from which unbiased sampling is not feasible.
In such situations, MapReduce and binning may be combined as a
pre-processing step for a wide range of statistical and scientific
analyses \citep{blocker2013}.

There are two common patterns for generating histograms of large data
sets in a single pass with MapReduce.  In the first method, each
mapper task generates a histogram over a subset of the data that it
has been assigned, serializes this histogram and sends it to one or
more reducer tasks which merge the intermediate histograms from the
mappers.

In the second method, illustrated in
Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
point to a bucket width and outputs that bucket as a key and '1' as a
value.  Reducers then sum up all of the values with the same key and
output to a data store.

>>>>

Protocol Buffers are used extensively at Google for almost all
RPC protocols, and to store structured information on a variety of
persistent storage systems \citep{dean2009designs}. Since the 
initial release in 2010, hundreds of Google's statisticians and
software engineers use the \pkg{RProtoBuf} package on daily basis 
to interact with these systems from within \proglang{R}.
The current section illustrates the power of Protocol Buffers to
collect and manage large structured data in one language 
before analyzing it in \proglang{R}. Our example uses MapReduce
\citep{dean2008mapreduce}, which has emerged in the last
decade as a popular design pattern to facilitate parallel 
processing of big data using distributed computing clusters.

Big data sets in fields such as particle physics and information
processing are often stored in binned (histogram) form in order 
to reduce storage requirements \citep{scott2009multivariate}. 
Because analysis over such large data sets may involve very rare
phenomenon or deal with highly skewed data sets or inflexible
raw data storage systems, unbiased sampling is often not feasible.
In these situations, MapReduce and binning may be combined as a
pre-processing step for a wide range of statistical and scientific
analyses \citep{blocker2013}.

There are two common patterns for generating histograms of large data
sets in a single pass with MapReduce. In the first method, each
mapper task generates a histogram over a subset of the data that it
has been assigned, serializes this histogram and sends it to one or
more reducer tasks which merge the intermediate histograms from the
mappers. In the second method, illustrated in
Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
point to a bucket width and outputs that bucket as a key and '1' as a
value.  Reducers then sum up all of the values with the same key and
output to a data store.
